Describe and Explore: Using feature-based time-series clustering to understand modern intensive longitudinal data (Working Title)

### 2022-11-27 manuscript.tex
:bug:
Updates from Overleaf

Once the important variables have been selected, the data needs to be prepared for the analysis steps. Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable. Two important steps are making the time-frames and response scales comparable across participants. Extracting features that describe the development of a month for one participant but 90 days for another participant might not be comparable for some psychological phenomena. This comparability is important on a conceptual level but the difference in data availability might also influence the clustering steps, where the participant with the shorter measurement period would, for example, have a larger number of missing values. It is thus generally advised to reduce the data to a common time frame. Once the time frames are comparable, also the responses should be comparable within and between participants. \hl{It is thus usually advised to standardize all variables for time series clustering.}

\Question{ \textit{Look up how to center/standardize (within- vs. between) and when (prior to extraction vs. prior to clustering)!}}

\subsection{Feature extraction}
aim: 
hundreds of possible features to describe psychological time series. One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series. However, such an approach is not always advisable for psychological time series. For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 180 time series features. More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results. This is particularly the case when features have a conceptual/psychological meaning. Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance). But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density). We would, thus, strongly advocate for a careful selection of time series features that are meaningful to the field and concepts.

@@

\section{Empirical Implementation/Illustration}
To illustrate the use case of the feature-based time series with psychological ESM data, we ....

@@

% Methods and Results from RMarkdown render
\input{methods-and-results}

>>>>

Once the important variables have been selected, the data needs to be prepared for the analysis steps. Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable. Two important steps are making the time-frames and response scales comparable across participants. Extracting features that describe the development of a month for one participant but 90 days for another participant might not be comparable for some psychological phenomena. This comparability is important on a conceptual level but the difference in data availability might also influence the clustering steps, where the participant with the shorter measurement period would, for example, have a larger number of missing values. It is thus generally advised to reduce the data to a common time frame.

\subsection{Feature extraction}
The main aim of feature extraction is 
hundreds of possible features to describe psychological time series. One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series. However, such an approach is not always advisable for psychological time series. For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 180 time series features. More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results. This is particularly the case when features have a conceptual/psychological meaning. Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance). But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density). We would, thus, strongly advocate for a careful selection of time series features that are meaningful to the field and concepts.

@@

\section{Empirical Illustration}
To illustrate the use case of the feature-based time series with psychological ESM data, we ....

@@

% Methods and Results from RMarkdown render
\input{methods-and-results}
\input{tables/descrWide.tex}
